---
title: 'cricketr flexes new muscles: The final analysis'
author: "Tinniam V Ganesh"
date: "March 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##
*’Twas brillig, and the slithy toves<br/> 
    Did gyre and gimble in the wabe: <br/>
  All mimsy were the borogoves, <br/>
    And the mome raths outgrabe' <br/>*
      
           Jabberwocky by Lewis Carroll
                       

No analysis of cricket is complete, without determining how players would perform in the host country.
Playing Test cricket on foreign pitches, in the host country, is a 'real test' for 
both batsmen and bowlers. Players, who can perform consistently both on domestic and foreign pitches
are the genuinely class players. Player performance on foreign pitches lets us differentiate the 
paper tigers, and home ground bullies among batsmen. Similarly, spinners who perform well, only on rank 
turners in home ground or pace bowlers who can only swing and generate bounce on specially prepared
pitches are not the genuine spinner or pace bowler.

So this post, helps in identifying those with real strengths and those who play good only when 
the conditions are in favor in home grunds. So this post brings a certain level of finality to
the analysis of players with my R package 'cricketr'

Besides, I also meant 'final analysis' in the literal sense, as I intend to take a long break
from cricket analytics and focus on some other domains like Neural Networks, Deep Learning and
Spark. 

As already mentioned, my R package 'cricketr' uses the statistics info available in ESPN Cricinfo Statsguru.
You should be able to install the package from CRAN and use  many of the functions available in the package. Please be mindful of  [ESPN Cricinfo Terms of Use](http://www.espncricinfo.com/ci/content/site/company/terms_use.html)

(Note: This page is also hosted as a GitHub page at cricketr and also at RPubs as cricketr: A R package for analyzing performances of cricketers

The data for the following players have already been obtained with the getPlayerData() function and
saved as *.csv files. I will be resuing these files, instead of getting them all over again. Hence the lines
have been commented.

For e,

The following lines gets player's data of matches played in England and against England
```{r }
#tendulkarEng=getPlayerData(35320,opposition=1,host=1,file="tendulkarVsEngInEng.csv",type="batting")
#laraEng=getPlayerData(52337,opposition=1,host=1,file="laraVsEngInEng.csv",type="batting")
#pontingEng=getPlayerData(7133,opposition=1,host=1,file="pontingVsEngInEng.csv",type="batting")
#sangakkaraEng=getPlayerData(50710,opposition=1,host=1,file="sangakkaraVsEngInEng.csv",type="batting")
```


```{r }
#tendulkarAus=getPlayerData(35320,opposition=2,host=2,file="tendulkarVsAusInAus.csv",type="batting")
#laraAus=getPlayerData(52337,opposition=2,host=2,file="laraVAusInAus.csv",type="batting")
#sangakkaraAus=getPlayerData(50710,opposition=2,host=2,file="sangakkaraVsAusInAus.csv",type="batting")
#kallisAus=getPlayerData(45789,opposition=2,host=2,file="kallisVsAusInAus.csv",type="batting")
#anCookAus=getPlayerData(11728,opposition=2,host=2,file="ancookVsWIInWI.csv",type="batting")

```

# Indian batsmen against Australia

```{r}
#tendulkarAus=getPlayerData(35320,opposition=2,host=2,file="tendulkarVsAusInAus.csv",type="batting")
#gavaskarAus=getPlayerData(28794,opposition=2,host=2,file="gavaskarVsAusInAus.csv",type="batting")
#kohliAus=getPlayerData(253802,opposition=2,host=12file="kohliVsEngInAus.csv",type="batting")
#sehwagAus=getPlayerData(35263,opposition=2,host=2,file="sehwagVsAusInAus.csv",type="batting")
```


#Indian batsmen against England

```{r}
#tendulkarEng=getPlayerData(35320,opposition=1,host=1,file="tendulkarVsEngInEng.csv",type="batting")
#dravidEng=getPlayerData(28114,opposition=1,host=1,file="dravidVsEngInEng.csv",type="batting")
#kohliEng=getPlayerData(253802,opposition=1,host=1,file="kohliVsEngInEng.csv",type="batting")
#gangulyEng=getPlayerData(28779,opposition=1,host=1,file="gangulyVsEngInEng.csv",type="batting")

```


#tendulkarEng=getPlayerData(35320,opposition=1,host=1,file="tendulkarVsEngInEng.csv",type="batting")
#dravidEng=getPlayerData(28114,opposition=1,host=1,file="dravidVsEngInEng.csv",type="batting")
#gavaskarEng=getPlayerData(28794,opposition=1,host=1,file="gavaskarVsEngInEng.csv",type="batting")
#kohliEng=getPlayerData(253802,opposition=1,host=1,file="kohliVsEngInEng.csv",type="batting")
#gangulyEng=getPlayerData(28779,opposition=1,host=1,file="gangulyVsEngInEng.csv",type="batting")


```{r}

#laraEng=getPlayerData(52337,opposition=1,host=1,file="laraVsEngInEng.csv",type="batting")
#pontingEng=getPlayerData(7133,opposition=1,host=1,file="pontingVsEngInEng.csv",type="batting")
#sangakkaraEng=getPlayerData(50710,opposition=1,host=1,file="sangakkaraVsEngInEng.csv",type="batting")
#ancookWI=getPlayerData(11728,opposition=4,host=4,file="ancookVsWIInWI.csv",type="batting")
```

You can also embed plots, for example:

```{r }
#steynInd=getPlayerData(47492 ,opposition=6,host=6,file="steynVsIndInInd.csv",type="bowling")
#mcgrathSL=getPlayerData(6565 ,opposition=8,host=8,file="mcgrathVsSLinSL.csv",type="bowling")
#kumble=getPlayerData(30176  ,opposition=1,host=1,file="kumbleVsEngInEng.csv",type="bowling")
#murali=getPlayerData(49636  ,opposition=1,host=1,file="muraliVsEngInEng.csv",type="bowling")
#warne=getPlayerData(8166  ,opposition=1,host=1,file="warneVsEngInEng.csv",type="bowling")

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
